---
title: "Capstone: Data Wrangling with Bank Data"
input_file: "bank-full.xlsx"
output_file: "bank_full_clean.csv"
---

##0: Load the data in RStudio and View it

```{r}
library(readxl)
bank_full <- read_excel("C:Users/.../Desktop/bank-full.xlsx")
View(bank_full)
```

##1: Clean up column names to be simple, short and descriptive
```{r}
str(bank_full)
#create a list of columns to change and names to change to
library(data.table)

org <- c("default","balance","housing","loan","contact","day","month","campaign","pdays","previous","poutcome","y")

chg <- c("default_credit","annual_balance","housing_loan","personal_loan","contact_type","last_contact_day","last_contact_month","contact_times","past_days","previous_contact_times","prev_outcome","outcome_term_deposit")

setnames(bank_full, old = org, new = chg)
```

##2:Check for missing values and display all unique values in each column
```{r}
colSums(is.na(bank_full))
lapply(bank_full,unique)
#since 573 entries in annual_balance and duration were omitted in results, we can summarize to see how the numbers range and what the average and median are. 
summary(bank_full$annual_balance)
summary(bank_full$duration)
```
##4: To rule out outliers,it'd be best to display plots of each variable to see which outliers should be removed from the dataset
```{r}
#numeric
library(ggplot2)
lapply(c("job", "marital", "education", "contact_type","prev_outcome"),
  function(col) {
    ggplot(bank_full, aes_string(col)) + geom_bar() + coord_flip()
  })
#categorical
lapply(c("age", "contact_times"),
  function(x) {
    hist(bank_full[[x]],xlab = x, main = x)
  })

#based on the info displayed in the plots, we should remove outliers for age which is below 20 
and above 80. Contact_times that exceed 10 times should also be removed from the data set rows.
Unknown job positions can be removed as well

library(dplyr)
bank_full <- bank_full %>% 
  filter(age>=20 & age<=80 & contact_times<=10, job!='unknown')

```



##5: Create dummy variables for columns that display binary results
```{r}

#return numeric value for TRUE FALSE
dummy_var <- function(x) {
  ifelse((x == "yes"),1,0)
}

dummy_col <- c("default_credit", "housing_loan", "personal_loan", "outcome_term_deposit")

for(i in dummy_col) {
  column_n <- paste("dummy", i, sep = "_")
  bank_full[[column_n]] <- dummy_var(bank_full[i])
}

str(bank_full)
```

##6: Submit the project on Github
```{r}
#Export results to csv file
write.csv(bank_full,"bank_full_clean.csv")
```



